Stack Safety as a Security Property

### 2020-07-09 main.tex
Fixes to properties I noticed while doing coq

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
What is stack safety? Numerous mechanisms have been proposed to guard
against attacks that corrupt the call stacks of running programs,
\ifaftersubmission
and many
others rely on a protected stack as an assumption to provide higher level
guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}
\bcp{Let's delete for now; later, we can try to incorporate it in the body
  of the paper.}
\fi,
but a clear and precise definition of stack safety is missing
from the literature.
%
We propose a formal definition phrased in the technical
terminology of language-based security---as a combination of a
confidentiality property (callees should not be able to read their caller's
stack frames) and an integrity property (callees should not corrupt their
caller's stack frames or control state).
  %
We discuss how these properties are enforced or approximated by
existing stack protection mechanisms and describe some useful
stronger and weaker variants.  \ifaftersubmission\bcp{should we mention stktokens here?}
\leo{Not sure. We almost never cite stuff in the abstract, do we?}\fi
\end{abstract}

@@

The call stack is a perennial target for attacks
based on memory safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \ifaftersubmission\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}
\bcp{Sure, but later!}
\fi

@@

In summary, we offer the following contributions:
%\bcp{I would still like to
%  see these points phrased in ``contribution normal form'': say exactly what
%  we did for each one (it's a bit vague---''considers,'' etc.), and
%  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
% +1 on Andrew's rewrite
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A new {\em formal stack-safety property} capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property can be enforced by
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, which is
  based on hardware tagging~(\cref{sec:conservative}).
\item
  A {\em lockstep} variant of the stack-safety property that is also
  enforced by the Depth Isolation policy
  but is more efficient to test, plus a proof (in Coq) that this variant is
  stronger than the original~(\cref{sec:lockstep}).
\item
  A {\em lazy} variant of the stack-safety property that can be enforced by the
  more efficient Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, plus a proof (in Coq) that this
  variant is weaker than the original~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.

We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).

@@

\apt{Some more words needed here to smooth the technicalities.}
\sna{Attempt...}

@@

\sna{Interestingly, split out like this, we don't need to talk about
     contour here -- they are varied in the next property.}
\begin{itemize}
  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
    some final state \(\mach_r'\) and any changes from the initial states are
    matched between \(\mach_r\) and \(\mach_r'\). In addition their
    observations are similar.
    \[\begin{split}
      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
      & ~ (\mach_e(\component) \not = \mach_r(\component)
        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
      \mach_r(\component) = \mach_r'(\component))
    \end{split}\]
  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
  observations are similar.
    \[\begin{split}
      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
      & \neg (\exists mo . \last(\machT) = mo)
        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
    \end{split}\]
\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
  its observations are a prefix of those of \(\machT\).
    \[\begin{split}
      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
      & \forall \mach_r ~ \pol_r ~ \obs_r .
        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \neg ~ R ~ \mach_r \Rightarrow \\
      & \pi_o(\MPT) \lesssim \pi_o(\machT)
    \end{split}\]
\end{itemize}

@@

If, for all such pairs of traces, \(\propdef{TraceConf} ~
\ret{\mach_c} ~ \MPT ~ \machT\), then the system as a whole enjoys
stack confidentiality with respect to \(\callmap\).

\apt{oy. still needs work to make this intelligible}
\sna{Attempted. There's a tension between introducing all the variables
     and keeping a good flow that I'm not sure I'm managing here though.}

@@

      The \(\mathit{restore}\) function takes a contour and four states: the
      original call state, a variant call state, and corresponding return
      states, and returns another state. The resulting state restores those
      components that are high confidentiality to match the original call,
      unless they were changed during execution:
      \[\mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')
        (\component) =
        \begin{cases}
          \mach_c(\component) & \text{if } \contour(\component) = (\HIGHSEC,\_)
            \text{ and } \\
          & \mach_e(\component) = \mach_r(\component) \text{ and }
            \mach_e'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}\]

@@

      {\em Observable trace confidentiality} is the lazy analog to trace
      confidentiality, and reuses the cases in which the traces don't return.
      For a contout \(\contour\), return predicate \(R\), and a pair of traces
      \(\MPT\) and \(\machT\), we consider the return case, in which
      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
      to those of \(\machT\). In place of the requirement that \(\mach_r\)
      and \(\mach_r'\) agree on components, we have the following requirement.

      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
      \mathit{restore}(\contour,\mach_e,\mach_e',\mach_r,\mach_r')\).
      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
      observational prefix of the induced trace from \(\mach_r'\),
      reflecting the possibility of a fail-stop after return.

@@

      \[\begin{split}
        \propdef{ObsTraceConf} ~ \contour ~ R ~ \MPT ~ \machT \triangleq
        & \mathit{ObsReturnCase} ~ \contour ~ R ~ \MPT ~ \machT \land \\
        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT \\
      \end{split}\]

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\), take any variation \(\mach_e'
      \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT | \ret{\mach_c}\)
      and \(\mach_e' \hookrightarrow \machT | \ret{\mach_c}\). A system
      enjoys observable stack confidentiality if, for all such
      \(\MPT\) and \(\machT\), \(\propdef{ObsTraceConf} ~
      \mathit{Cof}(\mach_c,n) ~ \ret{\mach_c} ~ \MPT ~ \machT\).

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
What is stack safety? Numerous mechanisms have been proposed to guard
against attacks that corrupt the call stacks of running programs, and many
others rely on a protected stack as an assumption to provide higher level
guarantees\bcp{this is a good point, but we don't pick it up anywhere in the
  paper, do we?}\leo{No, should we delete this? it still makes sense in flow}, but a clear and precise definition of stack safety is missing
from the literature.

We propose a formal definition of stack safety phrased in the technical
terminology of language-based security---as a combination of a
confidentiality property (callees should not be able to read their caller's
stack frames) and an integrity property (callees should not corrupt their
caller's stack frames or control state).
  %
We discuss how these properties are enforced or approximated by  
existing stack protection mechanisms and describe some useful
provably stronger and weaker variants.  \bcp{should we mention stktokens here?}
\leo{Not sure. We almost never cite stuff in the abstract, do we?}
\end{abstract}

@@

The call stack is a perennial target for attacks
based on memory safety vulnerabilities, with potential consequences
ranging from leakage or corruption of private stack data to
control-flow hijacking. To detect or prevent such attacks, a menagerie of
software and hardware techniques for stack safety have been proposed,
%
including stack canaries~\citep{Cowan+98},
bounds checking~\citep{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\citep{Kuznetsov+14},
shadow stacks~\citep{Dang+15,Shanbhogue+19},
capabilities~\citep{Woodruff+14,Chisnall+15,Skorstengaard+19,Skorstengaard+19b},
and hardware tagging~\citep{DBLP:conf/sp/RoesslerD18}. \iflater\apt{Mostly from
  nick; there could be more}\leo{Do we want to hunt for more refs here?}\fi

@@

In summary, we offer the following contributions:
%\bcp{I would still like to
%  see these points phrased in ``contribution normal form'': say exactly what
%  we did for each one (it's a bit vague---''considers,'' etc.), and
%  explicitly say why it is (a) useful, (b) challenging, and (c) novel.}
% +1 on Andrew's rewrite
\begin{itemize}
\item
  % \leo{More? Extensional}\bcp{We might need to invent a word for   this---''quasi-extensional''?}
  A novel formal stack safety property capturing the
  intuitive concepts of strict local state encapsulation and well-bracketed
  control flow~(\cref{sec:lse-and-wbcf}) in terms of security concepts.
  % These properties   rule out common examples of stack memory abuse.
  This property is enforced by 
  the Depth Isolation policy of \citet{DBLP:conf/sp/RoesslerD18}, based on hardware tags~(\cref{sec:conservative}).
\item
  A lockstep stack safety property that is also enforced by the Depth Isolation policy,
  and is more efficient to test, which we prove (in Coq) is stronger than the basic property~(\cref{sec:lockstep}).
\item
  A novel lazy stack safety property that is enforced by the more efficient
  Lazy Tagging and Clearing policy
  of \citet{DBLP:conf/sp/RoesslerD18}, which we prove (in Coq) is weaker than the basic property~(\cref{sec:lazy}).
%  This policy does not enforce our strict
%  local state encapsulation property,
%  but a modified version of it using unique per-activation tags enforces
%  an (also provably) weaker lazy variant of the property, which still prohibits
%  the same intuitively stack unsafe behaviors.
%
% \apt{Awkward, but the truth I fear.} \bcp{I agree the phrasing is awkward,
%   but the point we can make is still strong: To make the property easier
%   to enforce, we can slightly weaken its guarantees; the weaker version is
%   a bit more complicated to state but still retains the pragmatic
%%   protections of the original.}\apt{To clarify: Nick's LT\&C policy uses
%   depth-based tags, which do not enforce even our lazy policy. So we need to
%   introduce a version of LT\&C that uses unique activation tags.}

  %\item As evidence that the property is not too strong\apt{But we add the lazy version because the strict version \emph{is} too strong, in some sense.}, we show that it is
%  realizible by
%  a previously proposed micropolicy enforcement
%  mechanism~(\Cref{sec:micropolicy}). \leo{And it says yes on a
%    substantial body of code}. We also present a variation of the
%  property + policy that is lazy, easier/more efficient to implement,
%  and provably weaker (\Cref{sec:lazy}).
%\item A provably stronger (inductive? step-by-step/stepwise) variant
%  of this property tuned for testing/verification.  We test the whole
%  thing with a cool framework. \bcp{Which section?}
\end{itemize}
\Cref{sec:relwork} discusses related work;
\cref{sec:future} sketches directions for future work.
%
We begin with a description of our threat model
(\cref{sec:threat}), a running example that we refer to throughout the paper
(\cref{sec:running-example}),  and some technical preliminaries
(\cref{sec:prelim}).

@@

@@

\begin{itemize}
  \item If \(\MPT\) has a final state \(\last(\MPT) = \mpostate{\mach_r}
    {\pol_r}{\obs_r}\) and \(R\ \mach_r\), then \(\machT\) is also finite with
    some final state \(\mach_r'\) and any changes from the initial states are
    matched between \(\mach_r\) and \(\mach_r'\). In addition their
    observations are similar.
    \[\begin{split}
      \mathit{ReturnCase} ~ R ~ \MPT ~ \machT \triangleq &
      \forall ~ \mach_r ~ \pol_r ~ \obs_r ~ \component . \last(\MPT) =
        \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \exists \mach_r' ~ \obs_r' . \last(\machT) = (\mach_r',\obs_r')
        \land \pi_o(\MPT) \eqsim \pi_o(\machT) \land \\
      & ~ (\mach_e(\component) \not = \mach_r(\component)
        \lor \mach_e'(\component) \not = \mach_r'(\component) \Rightarrow
      \mach_r(\component) = \mach_r'(\component))
    \end{split}\]
  \item If \(\MPT\) is infinite, then \(\machT\) is also infinite and their
  observations are similar.
    \[\begin{split}
      \mathit{InfCase} ~ R ~ \MPT ~ \machT \triangleq
      & \neg (\exists mpo . \last(\MPT) = mpo) \Rightarrow \\
      & \neg (\exists mo . \last(\machT) = mo)
        \land \pi_o(\MPT) \eqsim \pi_o(\machT)
    \end{split}\]
\item If \(\MPT\) ends prematurely without fulfilling \(R\), then
  its observations are a prefix of those of \(\machT\).
    \[\begin{split}
      \mathit{FailCase} ~ R ~ \MPT ~ \machT \triangleq
      & \forall \mach_r ~ \pol_r ~ \obs_r .
        \last(\MPT) = \mpostate{\mach_r}{\pol_r}{\obs_r} \Rightarrow \\
      & \neg ~ R ~ \mach_r \Rightarrow \\
      & \pi_o(\MPT) \lesssim \pi_o(\machT)
    \end{split}\]
\end{itemize}

@@

For all such pairs of traces, \(\propdef{TraceConf} ~
\ret{\mach_c} ~ \MPT_{\mathit{call}} ~ \machT{\mathit{call}}\) must hold.
Trace confidentiality must also hold on the trace as a whole, so that
unallocated memory is still secret. Let \(\mach_0'
\approx_{\mathit{Cof}(\mach_0),0} \mach_0\) and \(\mach_0' \hookrightarrow
\machT\). Then \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\) must hold.

If these conditions hold, then the system as a whole enjoys
stack confidentiality with respect to \(\callmap\).

@@

      The \(\mathit{restore}\) function takes four states: the
      original entry state, a variant entry state, and corresponding return
      states. It returns another state. The resulting state restores those
      components that varied between entries to match the original entry,
      unless they were changed during execution:
      \[\mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')
        (\component) =
        \begin{cases}
          \mach_e(\component) & \text{if } \mach_e(\component) \not =
                                \mach_e'(\component) \text { and } \\
          & \mach_e(\component) = \mach_r(\component) \text{ and }
            \mach_e'(\component) = \mach_r'(\component) \\
          \mach_r'(\component) & \text{else} \\
        \end{cases}\]

@@

      {\em Observable trace confidentiality} is the lazy analog to trace
      confidentiality, and reuses the cases in which the traces don't return.
      For a return predicate \(R\) and a pair of traces
      \(\MPT\) and \(\machT\), we consider the return case, in which
      \(\last(\MPT)\) is some \(\mpostate{\mach_r}{\pol_r}{\obs_r}\) and
      \(R ~ \mach_r\). Then \(\last(\machT)\) is some \((\mach_r',\obs_r')\)
      and \(R ~ \mach_r'\). As above, the observations of \(\MPT\) are similar
      to those of \(\machT\). In place of the requirement that \(\mach_r\)
      and \(\mach_r'\) agree on components, we have the following requirement.

      Finally, let \(\mach_e\) be the machine state of \(\head(\MPT)\),
      and \(\mach_e'\) be that of \(\head(\machT)\). We restore \(\mach_r'\)
      based on \(\mach_e\) and \(\mach_e'\) to get \(\mach_r'' =
      \mathit{restore}(\mach_e,\mach_e',\mach_r,\mach_r')\).
      The induced trace from \(\mpstate{\mach_r}{\pol_r}\) should be an
      observational prefix of the induced trace from \(\mach_r'\),
      reflecting the possibility of a fail-stop after return.

@@

      \[\begin{split}
        \propdef{ObsTraceConf} ~  R ~ \MPT ~ \machT \triangleq
        & \mathit{ObsReturnCase} ~ R ~ \MPT ~ \machT \land \\
        & \mathit{InfCase} ~ R ~ \MPT ~ \machT \land \mathit{FailCase} ~ R ~
          \MPT ~ \machT \\
      \end{split}\]

@@

      {\em Observable stack confidentiality} is structured like stack
      confidentiality. Let \(\callmap\) be a callmap; for each initial
      state \(\mach_0\) and \(n\)-transition from \(\mpstate{\mach_c}{\pol_c}\)
      to \(\mpstate{\mach_e}{\pol_e}\) in the induced trace
      \(\mpstate{\mach_0}{\pinit(\mach_0)} \hookrightarrow \MPT\), take any
      variation \(\mach_e' \approx_{\mathit{Cof}(\mach_c,n)} \mach_e\). Let
      \(\mpstate{\mach_e}{\pol_e} \hookrightarrow \MPT_{\mathit{call}} |
      \ret{\mach_c}\) and \(\mach_e' \hookrightarrow \machT_{call} |
      \ret{\mach_c}\). We must have \(\propdef{ObsTraceConf} ~
      \ret{\mach_c} ~ \MPT ~ \machT\).

      Additional, for any variation \(\mach_0' \approx_{\mathit{Cof}(\mach_0),0}
      \mach_0\), taking the induced trace  \(\mach_0' \hookrightarrow
      \machT\), we must have \(\propdef{TraceConf} ~ \emptyset ~ \MPT ~ \machT\).

      A system enjoys observable stack confidentiality if observable
      confidentiality holds in all such cases.
